---
title: "Penguins Assignment Question 4"
date: "2022-12-06"
candidate number: 1054368
output: 
 pdf_document: default
 html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


*Loading the packages*
```{r}
library(ggplot2)
library(palmerpenguins)
suppressPackageStartupMessages(library(janitor))
suppressPackageStartupMessages(library(dplyr))
library(tidyr)
library(ragg)

```

```{r}
#Setting the working directory 
setwd("~/Biology/Year 3/Year 3 R/Computing Assessments/PenguinProject") 
```


**Question 4: Run a statistical test on the Palmer Penguins dataset and produce a figure to explain it**

**Cleaning the Data**

```{r}
penguins_raw
#want to save this penguins_raw dataset
write.csv(penguins_raw, paste0("data_raw/penguins_raw.csv"))
#Saved penguins raw to a data raw folder

#Now we are defining the cleaning function and what it will do to penguins_raw 
cleaning <- function(penguins_raw) {
  penguins_raw %>% 
    select(-starts_with("delta")) %>%
    select(-Comments)%>%
    clean_names()}

#This is within the cleaning.r function that is saved separately. It removes NA
#values and selects the variables I am including in my data analysis
remove_empty_mass_and_length <- function(penguins_clean){
    penguins_clean %>%
    filter(!is.na(culmen_length_mm)) %>%
    filter(!is.na(body_mass_g)) %>%
    select(body_mass_g, culmen_length_mm, species)
}

source("functions/cleaning.r")
#This specifies where the cleaning function is saved

#We are then applying this function to the penguins_raw dataset
penguins_clean <- cleaning(penguins_raw)
penguins_now_clean <- remove_empty_mass_and_length(penguins_clean)

#Saving penguins clean dataset 
write.csv(penguins_clean, paste0("data_clean/penguins_clean.csv"))

#Saving penguins now clean dataset
write.csv(penguins_now_clean, paste0("data_clean/penguins_now_clean.csv"))

```

Now the data has been cleaned, I am going to test to see if there is a significant correlation between body mass and culmen length. I will use Pearson's correlation coefficient because the two variables are numerical and continuous 

H0 -> There is no significant correlation between culmen length and body mass in Palmer Penguins

HA -> There is a significant correlation between culmen length and body mass in Palmer Penguins

**Testing Assumptions**

```{r}
penguins_model <- lm(culmen_length_mm ~ body_mass_g, data=penguins_now_clean) 
#Making a linear regression model for body mass and culmen length 

summary(penguins_model) #Summary of the model 

#Checking the normality assumptions for the regression model
plot(penguins_model, which =2)#QQplot
plot(penguins_model, which =1)#Residuals vs Fitted
#The points adhere well to the line in the qqplot, and the red line lies close to 0 
#throughout the residuals vs fitted plot
#So the assumptions of normality and homogeneity of variance are well-met for the model


#Can do Pearson's correlation coefficient if the assumption of bivariate normality is not breached. 
#I will make histograms for each variable to see if this is the case 
ggplot(data=penguins_now_clean, 
       aes(x=body_mass_g))+geom_histogram(colour="black",fill="blue")

ggplot(data=penguins_now_clean,
       aes(x=culmen_length_mm))+ geom_histogram(colour="black",fill="orange")

```

From these histograms, it appears that the assumption of bivariate normality is not well-met, since the distribution for body mass is skewed to the left, whereas the culmen length distribution does not have this same skew. Therefore, I will use a non-parametric test to test for a significant correlation. Instead of Pearson's, I will use Spearman's Correlation Coefficient. 

**Statistical Test and Graph Plotting**

```{r}

#Performing the Spearman's correlation coefficient
cor.test(penguins_clean$body_mass_g, penguins_clean$culmen_length_mm, 
         method=c("spearman"),exact=FALSE)
#That has given me a p value of 2.2e-16. Suggests that there is a significant correlation 
#between the two variables. Will now visualise this with a scatter plot


#Defining my functions for plotting a graph for all the penguins, and broken down by penguin species
plot_penguin_graph1 <- function(penguins_now_clean){
  penguins_now_clean %>%
    ggplot(
           aes(x=body_mass_g, y=culmen_length_mm))+geom_point()+
    geom_smooth(method="lm", colour="blue")+labs(x="Body Mass (g)", y="Culmen Length (mm)", 
    title="Correlation Between Body Mass and Culmen Length 1")+
    theme_bw()
}   
    
plot_penguin_graph2 <- function(penguins_now_clean){ 
  penguins_now_clean %>%
    ggplot(
           aes(x=body_mass_g, y=culmen_length_mm, colour=species))+ geom_point()+
    scale_colour_manual(values=c("deeppink", "blue", "orange"))+ geom_smooth(method="lm")+
    labs(x="Body Mass (g)", y="Culmen Length (mm)", 
    title="Correlation Between Body Mass and Culmen Length 2", colour="Penguin Species") 
}


source("functions/plotting.r")
#This is where my function is saved

#Calling the plotting function to make penguin_graph1 and penguin_graph2
penguin_graph1 <- plot_penguin_graph1(penguins_now_clean) 
penguin_graph2 <- plot_penguin_graph2(penguins_now_clean)

#Looking at the graphs
penguin_graph1 
penguin_graph2

```

The output of the Spearman's rank gave a P value of <2.2e-16. This is less than 0.05, so we can reject the null hypothesis that there is no significant correlation between body mass and culmen length. As for the strength of the correlation, the rho value is 0.584, representing a strong positive correlation. 

When this result is visualised with a scatterplot and regression lines, we see that there is a positive correlation between body mass and culmen length such that larger penguins have a longer culmen. This means that body mass is a good predictor of culmen length. The second scatterplot breaks down this positive correlation for each species, so we can see that the correlation holds true across all species of penguins.This is important to visualise to avoid an undetected Simpson's Paradox, where an overall trend hides variations or deviations from this trend within each group. In this case, there is no Simpson's Paradox because there is still a positive correlation for all the penguin species. 

**Saving the figures**

```{r}
#Defining a function for saving penguin graph 1 and penguin graph 2

save_graph1_png <- function(penguins_now_clean,
                                  filename, size, res, scaling){
  agg_png("figures/penguin_graph1.png", 
                    width   =  25, 
                    height  =  20, 
                    units   =  "cm", 
                    res     =  600, 
                    scaling =  1)
  penguin_graph1 <- plot_penguin_graph1(penguins_now_clean)
  print(penguin_graph1)
  dev.off()
}

save_graph2_png <- function(penguins_now_clean,
                                  filename, size, res, scaling){
  agg_png("figures/penguin_graph2.png", 
                    width   =  25, 
                    height  =  20, 
                    units   =  "cm", 
                    res     =  600, 
                    scaling =  1)
  penguin_graph2 <- plot_penguin_graph2(penguins_now_clean)
  print(penguin_graph2)
  dev.off()
}

source("functions/saving.r") 
#this is where my saving function is stored

#Calling the saving function to save penguin graph 1 and penguin graph 2 as png files
penguin_graph1.png <- save_graph1_png(penguins_now_clean)
penguin_graph2.png <- save_graph2_png(penguins_now_clean)

#my graphs have now been saved in the "figures" folder 

```


